---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Course Project 1

## Loading and preprocessing the data

First the data is load and treated

```{r load, warning=FALSE,message=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
data<-read.csv("activity.csv")

data <-data %>%
        mutate(steps=steps,
               time=as.POSIXct(
                       ymd(as.character(date))+minutes(interval)))

```


## What is mean total number of steps taken per day?

Then, mean and median of the daily sum are calculated. NAs are disconsidered.

```{r stepsPerDay}

daily<-data%>%
        group_by(date)%>%
        summarize(steps=sum(steps, na.rm=T))


dailymean<-mean(daily$steps)
dailymedian<-median(daily$steps)

p<-ggplot(data=daily)
p+geom_histogram(aes(steps))+
        geom_vline(aes(xintercept=dailymedian,color='median') ,size=1.1) +
        geom_vline(aes(xintercept=dailymean,color='mean'),size=1.1)+
        scale_color_manual(name = "statistics",
                           values = c(median = "red", mean = "blue"))

```

The mean is `r dailymean` and the median is `r dailymedian` 

## What is the average daily activity pattern?



Then, a plot of the timeseries of the avarage value for each interval is calculated
and the timeseries for an avarage day is plotted.


```{r avarageday}

avarageday<-data%>%
        group_by(interval)%>%
        summarize(steps=mean(steps, na.rm=T))


maximum<-data$interval[which(avarageday$steps==max(avarageday$steps))]

p<-ggplot(data=avarageday,aes(interval,steps))
p+ geom_line() 


```

The interval with more steps is the one starting at `r maximum` minutes.



## Imputing missing values

Then, mean and median of the daily sum are calculated. NAs values weree substituted 
for the values of an avarage day.

```{r imputmissing}
totNAs<-sum(is.na(data$steps))

data2<-data%>%mutate(steps=ifelse(is.na(steps),round(
        avarageday$steps[interval/5+1]),steps))

daily<-data2%>%
        group_by(date)%>%
        summarize(steps=sum(steps, na.rm=T))


dailymean<-mean(daily$steps)
dailymedian<-median(daily$steps)

p<-ggplot(data=daily)
p+geom_histogram(aes(steps))+
        geom_vline(aes(xintercept=dailymedian,color='median') ,size=1.1) +
        geom_vline(aes(xintercept=dailymean,color='mean'),size=1.1)+
        scale_color_manual(name = "statistics",
                           values = c(median = "red", mean = "blue"))

```

The mean is `r dailymean` and the median is `r dailymedian`. Considerably bigger than in
the previous exercise.


## Are there differences in activity patterns between weekdays and weekends?

Then, the data is stratified into weekends and weekdays. And timeseries are ploted for these 
two cases. It can be verified very different patterns

```{r weekend}
library(chron)

data<-data %>%
        mutate(weekend=is.weekend(time))

avarageday<-data%>%
        group_by(interval,weekend)%>%
        summarize(steps=mean(steps, na.rm=T))

p<-ggplot(data=avarageday,aes(interval,steps))
p+ geom_line()+facet_grid(weekend~.)

```
