---
title: "CourseProject1"
author: "André Costa"
date: "10 de abril de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Course Project 1

First the data is load and treated

```{r load, warning=FALSE,message=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
data<-read.csv("activity.csv")

data <-data %>%
        mutate(steps=steps,
               time=as.POSIXct(
                       ymd(as.character(date))+minutes(interval)))

```

Then, mean and median of the daily sum are calculated. NAs are disconsidered.

```{r stepsPerDay}

daily<-data%>%
        group_by(date)%>%
        summarize(steps=sum(steps, na.rm=T))


dailymean<-mean(daily$steps)
dailymedian<-median(daily$steps)

p<-ggplot(data=daily)
p+geom_histogram(aes(steps))+
        geom_vline(aes(xintercept=dailymedian,color='median') ,size=1.1) +
        geom_vline(aes(xintercept=dailymean,color='mean'),size=1.1)+
        scale_color_manual(name = "statistics",
                           values = c(median = "red", mean = "blue"))

```

The mean is `r dailymean` and the median is `r dailymedian` 

Then, a plot of the timeseries of the avarage value for each interval is calculated
and the timeseries for an avarage day is plotted.


```{r avarageday}

avarageday<-data%>%
        group_by(interval)%>%
        summarize(steps=mean(steps, na.rm=T))


maximum<-data$interval[which(avarageday$steps==max(avarageday$steps))]

p<-ggplot(data=avarageday,aes(interval,steps))
p+ geom_line() 


```

The interval with more steps is the one starting at `r maximum` minutes.

Then, mean and median of the daily sum are calculated. NAs values weree substituted 
for the values of an avarage day.

```{r imputmissing}
totNAs<-sum(is.na(data$steps))

data2<-data%>%mutate(steps=ifelse(is.na(steps),round(
        avarageday$steps[interval/5+1]),steps))


maximum<-data$interval[which(daily$steps==max(daily$steps))]

p<-ggplot(data=avarageday,aes(interval,steps))
p+ geom_line() 


daily<-data2%>%
        group_by(date)%>%
        summarize(steps=sum(steps, na.rm=T))


dailymean<-mean(daily$steps)
dailymedian<-median(daily$steps)

p<-ggplot(data=daily)
p+geom_histogram(aes(steps))+
        geom_vline(aes(xintercept=dailymedian,color='median') ,size=1.1) +
        geom_vline(aes(xintercept=dailymean,color='mean'),size=1.1)+
        scale_color_manual(name = "statistics",
                           values = c(median = "red", mean = "blue"))

```

The mean is `r dailymean` and the median is `r dailymedian`. Considerably bigger than in
the previous exercise.

Then, the data is stratified into weekends and weekdays. And timeseries are ploted for these 
two cases.

```{r weekend}
library(chron)

data<-data %>%
        mutate(weekend=is.weekend(time))

avarageday<-data%>%
        group_by(interval,weekend)%>%
        summarize(steps=mean(steps, na.rm=T))

p<-ggplot(data=avarageday,aes(interval,steps))
p+ geom_line(aes(color=weekend))

```
